Verifying Concurrent Search Structure Templates for C Programs using VST

### 2023-04-16 concurrency/CONCUR 2023/paper.tex
WIP give-up template

$$\forall \texttt{t}.\ \left\langle  \infp (\texttt{pn->n}) \ | \ \treerep\ \texttt{t} \right\rangle$$ 
$$\texttt{traverse(pn, k)}$$
$$\left\langle \texttt{res} \ \texttt{n'}. \ (\texttt{res} \leftrightarrow (\texttt{pn->n'->t} = \texttt{NULL})) \ast  \infp (\texttt{pn->n'}) \ast \texttt{R(pn->n')} \ast (\texttt{k} \in \texttt{range})
\ |\ \treerep\ \texttt{t} \right\rangle$$

In contrast to the more intricate specification of the lock-coupling style \texttt{traverse} function, the give-up template \texttt{traverse} function's specification is comparatively simpler. The local pre-condition for the specification solely comprises $\infp (\texttt{pn->n})$, a footprint predicate signifying that node \texttt{n} resides within the search structure's footprint. Subsequently, the \texttt{traverse} function operates on the entire search structure, returning a subsequent node \texttt{n'}, where \texttt{k} lies within the node \texttt{n'}'s range. As a result, the local post-condition is anticipated to assert that \texttt{n'} is also positioned within the footprint predicate, the resource \texttt{R(pn->n')} embodies the concrete information concerning node \texttt{n'}, and a Boolean \texttt{res} indicates whether \texttt{pn->n'->t} is \texttt{NULL} or not.

>>>>

$$\forall \texttt{t}.\ \left\langle \infp (\texttt{pn->n}) \ | \ \treerep\ \texttt{t} \right\rangle$$ 
$$\texttt{traverse(pn, k)}$$
$$\left\langle \texttt{res} \ \texttt{n'}. \ (\texttt{res} \leftrightarrow (\texttt{pn->n'->t} = \texttt{NULL})) \ast  \infp (\texttt{pn->n'}) \ast \texttt{R(pn->n')} \ast (\texttt{k} \in \texttt{range(n')})
\ |\ \treerep\ \texttt{t} \right\rangle$$

In contrast to the more intricate specification of the lock-coupling style \texttt{traverse} function, the give-up template \texttt{traverse} function's specification is comparatively simpler. The local pre-condition for the specification solely comprises $\infp(\texttt{pn->n})$, a footprint predicate signifying that node \texttt{n} reside within the search structure's footprint. Subsequently, the \texttt{traverse} function operates on the entire search structure, returning a subsequent node \texttt{n'}, where \texttt{k} lies within the node \texttt{n'}'s range. As a result, the local post-condition is anticipated to assert that \texttt{n'} is also positioned within the footprint predicate, the resource \texttt{R(pn->n')} embodies the concrete information concerning node \texttt{n'}, and a Boolean \texttt{res} indicates whether \texttt{pn->n'->t} is \texttt{NULL} or not.

We define the loop invariant for the give-up template of \texttt{traverse} function as follows: 
\begin{align*} &\texttt{traverse\_inv(pn)} \triangleq\ \exists \texttt{n'}.\ \infp (\texttt{pn->n}) \ast \infp (\texttt{pn->n'})   
\end{align*}

The proof starts with $\infp (\texttt{pn->n})$ predicate from the local pre-condition. We get the resource $\texttt{R(pn->n)}$ into the thread-local state after acquiring the lock of the node $\texttt{pn->n}$ (line 5 in Figure \ref{traverse_giveup_b}). In the \texttt{then} of \texttt{inRange}, this give us the additional predicate $\texttt{k} \in \texttt{range(n)}$. Similar to proof of \texttt{traverse} of lock-coupling template, it stops when it reaches the base case, which occurs when the current node is \lstinline{NULL} (see line 8 in Figure \ref{traverse_giveup_b}). At this point, the specification indicates that we have $\texttt{k} \in \texttt{range(n')} \ast \texttt{pn->n'->t} = \texttt{NULL}$, enabling us to prove the post-condition. On the other hand, the function terminates if \lstinline{findNext} is unable to locate a subsequent node with the key value  $\texttt{k}$ within its set of key values. Otherwise, the \lstinline{findNext} successfully detects the next node (for example, \texttt{n'}) to visit, then it give a predicate that asserts \texttt{n'} is in the footprint of structure, $\infp(\texttt{pn->n})$. From that, after releasing the lock of node the current node \texttt{p} (since \texttt{p} and  \texttt{n} point to the same memory address - line 6 in Figure \ref{traverse_giveup_b}), we can prove it satisfies the loop invariant $\texttt{traverse\_inv(pn)}$ and complete this branch.

In the \texttt{else} branch of the call to \texttt{inRange}, since \texttt{k} is not in the range of the current node \texttt{p}, it gives up, and relinguishes the lock on \texttt{p}, the goes back to the root of data structure to re-try.
